Experience-Biased Symbolic Planning

### 2014-05-16 AAAI_AuthorKit/paper.tex
SoCS 2014 submission

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
% AAAI format packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
% Additional packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{comment}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
% END Additional packages
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Experience-Biased Symbolic Planning)
/Author (Aram Ebtekar, Mike Phillips, Sven Koenig, Maxim Likhachev)
/Keywords (symbolic planning, E-graph, experience-biasedh, A* search, STRIPS, HSP)
}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Experience-Biased Symbolic Planning}
\author{Aram Ebtekar$^\dagger$ \and Mike Phillips$^\dagger$ \and Sven Koenig\thanks{University of Southern California, Los Angeles, CA 90089} \and Maxim Likhachev% <-this % stops a space
\thanks{Carnegie Mellon University, Pittsburgh, PA 15217}% <-this % stops a space
%
}
\maketitle
\begin{abstract}
\begin{quote}
Weighted A* search is the foundation for state-of-the-art domain-independent symbolic planners.
An important open challenge for these planners is to utilize experience with similar planning problems to speed up subsequent planning episodes.
Experience graphs have recently shown promise as a technique for plan reuse in the context of weighted A* searches in robotics, by biasing them towards a subgraph of the search space, typically chosen to consist of the edges used in previous plans.
In this paper, we demonstrate how to augment symbolic planners with
experience graphs, using the HSP2 planner as a representative
example.
The resulting planners are able to trade off seamlessly
between plan quality and runtime and thus have advantages over past
plan-reuse techniques, such as planning by analogy.
We demonstrate experimentally that on many symbolic domains, our approach reduces the number of nodes that the HSP2 planner needs to consider by a factor of 2 or more while achieving about the same plan quality.
\end{quote}
\end{abstract}

@@

In practice, even using state-of-the-art methods to generate heuristics on isolated planning instances, STRIPS problems can be very difficult to solve without expert domain knowledge.
Even humans have difficulty when faced with an unfamiliar kind of puzzle, though we get better with experience.
Thus, one might hope that a planning agent would similarly learn to generalize solutions from past planning experiences to related new instances.

\end{comment}

In forward weighted A* search, a frontier of nodes expands from the start until it hits a goal, at which point a solution is found. \textbf{Heuristics} are cheaply computable approximations of the remaining distance from a goal to the node. A* uses these to focus the search toward the goal.

However, without domain expert knowledge, symbolic planning problems can be very difficult. A state-of-the-art domain-independent planner, or even a human faced with an unfamiliar kind of puzzle, might have great difficulty. Like a human, then, one might hope that a planning agent would similarly learn to generalize solutions from past planning experiences to related new instances.

A recent approach builds \textbf{experience graphs} (E-graphs) to estimate the high-level connectivity of the free space in motion planning tasks \cite{phillips2012graphs}.
The intuition is to remember previously generated paths so that, when a new start and goals are queried, a new path can be quickly generated by reusing subpaths from the E-graph.
A variant of weighted A* is employed which biases its focus toward the E-graph.
While this work demonstrates promising results in robot motion planning, E-graphs have never before been applied to symbolic domain representations such as the STRIPS language.

Our present contribution is to investigate the use of E-graphs in STRIPS domains. In particular, we take the HSP2 planner \cite{bonet2001planning} as a representative of the class of modern weighted A* based symbolic planners, and augment its heuristic with a bias toward its E-graph. After presenting the related works, we begin by describing the STRIPS language and the original HSP heuristic. In the following section, we formally introduce E-graphs in technical detail. Then we combine the two approaches, resulting in the new Experience-Biased HSP planner. We present the algorithm and discuss its theoretical properties. This is followed by experiments showing promise for the application of E-graphs to symbolic problem solving. Finally, we conclude with some thoughts on extensions worth investigating.

@@

Classically, case-based reasoning methods 
such as PRIAR~\cite{Kamb:92}, SPA~\cite{Hank:95}, MRL~\cite{Koeh:94}, and NoLimit~\cite{mljournal}
have been used to recall 
previous plans that appear similar to the current problem and 
then adapt them to meet the new specifications.
Compared to these methods, our work reuses prior experience in
the context of modern weighted A* based methods for symbolic planning. We also provide
a bound on the suboptimality of the solution.

Recently there has been work on incorporating prior knowledge 
into modern planners (based on weighted A* and powerful heuristics).
For instance, GPG~\cite{Gere:00} uses planning graphs to adapt a prior path to the 
new scenario by searching locally within windows around plan 
inconsistencies. To guarantee completeness, the windows can grow 
if solutions are not found until they contrain the whole plan.
In this work~\cite{DBLP:conf/iccbr/RosaOB07} case-based reasoning is used to choose the 
order of state evaluations of EHC (enforced hill-climbing) which 
is a component in modern planners such as FF.

OAKplan~\cite{Serina:2010:KFC:1860143.1860472} is a recent 
case-based reasoning method which converts 
planning instances into a compact graph representation and 
then uses recent kernel methods for matching labeled graphs to select
a similar case to the problem at hand. The selected case is then 
adapted using LPG-ADAPT~\cite{Fox06planstability} LPG-ADAPT is a replanning method
that encourages \textit{plan stability}, which tries to make 
sure that the new plan is as similar to the previous plan as possible. 
The method also uses a modern heuristic search approach.

ERRT-PLAN~\cite{workshop-icaps12-errtplan} presents a very different 
approach to reuse in planning. 
Similar to E-Graphs, ERRT has its roots in motion planning~\cite{Bruce:2002}.
ERRT-PLAN applies this sampling-based motion planner that leverages 
reuse to symbolic planning. The planner randomly chooses to extend 
the search tree toward the goal, toward goals from previous 
plans, or to reuse an action from a previous plan. 
The extension operations are based on modern heuristic search
planners.
The stochastic nature of the direction the search progresses toward
allows it to overcome local minima caused by misleading heuristics.

The primary difference between our approach and previous ones 
is that we plan with prior experience in a way that have provable
bounds on suboptimality of the solution. The use of E-Graphs also
allows us to formulate this in a way that the user has control over
this suboptimality bound with an intuitive parameter that trades
reuse (speed) for solution quality.

The HSP2 \cite{bonet2001planning} planner uses the heuristic pioneered by HSP, but replaces greedy hill-climbing with weighted A* search.
We test our approach on HSP2 because the latter is straightforward to describe and it is representative of the core techniques that are common to many state-of-the-art domain-independent symbolic planners.
Some newer additions to the field include the causal graph heuristic of the Fast Downward planner \cite{helmert2006fast} and the landmark heuristic of the LAMA planner \cite{richter2010lama}. Our approach is orthogonal to these developments and can, in principle, be adapted to work with any search heuristic.

@@

Many of today's state-of-the-art domain-independent STRIPS planners are based on the HSP2 planner \cite{bonet2001planning}, which we now describe.
It's common to compute heuristics by solving an easier, relaxed version of the original problem. In STRIPS, one might ignore the delete lists of operations.
Since having more atoms makes preconditions and the goal condition more likely to hold, this can only make the problem easier, so the resulting heuristic never overestimates the true cost.

However, even the relaxed STRIPS planning problem includes set-cover as a special case, making it NP-hard.
Intuitively, we see that the relaxed search space remains exponential.
To reduce it, HSP2 decouples the atoms, instead estimating the cost of achieving each individual atom.

Let's say we want a heuristic estimate of the distance from $S$ to a goal that contains $G$. HSP2 estimates the cost to achieve an atom $a\in A$ from $S$ by $g_S(a) = $
\[\begin{cases} 0  &\mbox{if } a \in S
\\ \min_{op\mid a\in Add(op)} \left(g_S(Prec(op)) + Cost(op)\right)  &\mbox{if } a \notin S \end{cases}\]

The HSP heuristic (also used in HSP2) is then defined by
\[h^{HSP}(S) = g_S(G),\]
the estimated cost of achieving all atoms in $G$. Note that $g_S(\cdot)$ is evaluated on certain atom sets, namely $Prec(op)$ and $G$.
To keep the computations feasible, define $g_S(P) = \max_{p\in P} g_S(p)$ for atom sets $P$.
The resulting HSP-max heuristic is an underestimate; indeed, it satisfies a stronger property called \textbf{consistency}.

\begin{defn} For $\epsilon\ge 1$, a heuristic $h$ is called \textbf{$\epsilon$-consistent} iff $h(S) = 0$ when $S$ is a goal and $h(S) \le \epsilon c(S,S') + h(S')$ for all $S,S'\in V$. $h$ is \textbf{consistent} if it is $1$-consistent. \end{defn}
\begin{thm} Given an $\epsilon$-consistent heuristic, A* without reexpansions is guaranteed to find a path costing no more than $\epsilon$ times the optimal path cost \cite{LikGorThr-ara}. \end{thm}

If the atoms were truly independent, a more accurate estimate would be $g_S(P) = \sum_{p\in P} g_S(p)$.
Despite its inconsistency in general, the latter HSP-plus heuristic is often useful in practice, as it uses more information and biases more greedily toward the goal.

@@

% TODO delete
In a sense, a lot of work is wasted by estimating the cost to reach every atom, when we only require $g_S(G)$.
Later, we'll see how to make better use of this computation.

@@

E-graphs can improve total search time over a series of related queries on the same graph.
Formally, an E-graph is a subgraph of the state space.
It is updated between queries, typically by adding the solution path from the previous search query.

Suppose we have a consistent heuristic $h(S,S')$ for the distance between arbitrary pairs of states.
That is, $h$ obeys the triangle inequality, is never dominated by an edge cost, and $h(S,S) = 0$ for all $S\in V$.
%TODO Planning with E-graph is simply running A* or WA* with heuristic h^E.
The E-graph heuristic $h^E$ biases the search to follow E-Graph edges instead of $h$ by penalizing the latter by an inflation factor $\epsilon^E > 1$.
To be precise, define

\[h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N \min \{\epsilon^E h(S_{i-1},S_i),c^E(S_{i-1},S_i)\}\]
with the minimization taking place over all goal states $S_N$ and paths $\pi = \langle S_0,S_1,...,S_N \rangle$. $c^E$ are E-graph edge costs, or $\infty$ if the E-graph edge does not exist.

In the limit as $\epsilon^E \rightarrow 1$, E-graph edges offer no advantage as $h(S_{i-1},S_i) \le c^E(S_{i-1},S_i)$. Hence, using the triangle inequality to coalesce the sum,
\[h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N h(S_{i-1},S_i) = \min_{S_N} h(S_0,S_N)\]

Conversely, as $\epsilon^E \rightarrow\infty$, the inflated terms dominate. Hence, the minimizing path becomes the one which incurs the least estimated cost outside the E-graph. In particular, $h^E$ has the following property:

\textit{For sufficiently large $\epsilon^E$, once the A* frontier touches an E-graph component which includes a goal state, no node outside the E-graph will ever again need to be expanded.}

This means, for instance, that if the E-graph already contains a path from the start to a goal, then A*, guided by $h^E$ with large bias parameter $\epsilon^E$, will never leave the E-graph.
If the E-graph is small, a solution would promptly be discovered.
In this manner, little work is needed to solve previously encountered subproblems.

We cannot afford to compute $h^E$ according to its literal definition, as there are far too many paths to consider.
Fortunately, there exists a practical means of computing it. By the triangle inequality, any consecutive pair of $h(S_{i-1},S_i)$ terms can be merged into one.
Thus, we lose no generality in restricting $S_i$ to lie on the E-graph for $0 < i < N$.
The heuristic computation can focus on the E-graph, using $h(S_{i-1},S_i)$ to estimate the cost of leaving the E-graph at $S_{i-1}$ before re-entering it at $S_i$.

Let $V^E$ be the union of the goal states and the E-graph's vertices.
In addition to following E-graph edges, we imagine it's possible to ``jump" from any $S\in V^E$ to any $S'\in V^E$ at cost $\epsilon^E h(S,S')$.
That is, let $c'(S,S') = \min\left(c^E(S,S'),~\epsilon^E h(S,S')\right)$

In a preprocessing stage before the main search, we apply Dijkstra's algorithm once in reverse with the costs $c'$ to compute the estimated distance-to-goal $h^E(S)$ from every $S\in V^E$. Ignoring the computation of $h$ (presently considered as a blackbox), this preprocessing takes $O(|V^E|^2)$ time, which is insignificant for small E-graphs and goal sets.

Later, when generating $S \notin V^E$, we see that we have already precomputed costs of paths consisting of all but the first edge of the minimizing path $\pi$ in the definition of $h^E(S)$. Thus, we derive the mathematically equivalent but much cheaper computation
\[h^E(S) = \min_{S'\in V^E} \left(\epsilon^E h(S,S') + h^E(S')\right)\]

\begin{thm} If the base heuristic $h$ is consistent, then $h^E$ is $\epsilon^E$-consistent. \cite{phillips2012graphs}\end{thm}

\begin{figure}
	\begin{center}
	\includegraphics[scale=0.5]{Pentagon.png}
	\end{center}
	\caption{An example E-graph and corresponding $h^E$-values.}
	\label{fig:example}
\end{figure}

Figure \ref{fig:example} illustrates the idea. Suppose the dark edges have cost 3 and make up the E-graph. Off-E-graph edges are not shown, as they are only accessed indirectly via $h$; instead, light edges represent the jump estimates $h(S_i,S_j) = 2$ (with one exception: $h(S,G) = 5$), which we inflate by a factor $\epsilon^E=2$ to make 4 (or 10, respectively). Dijkstra's algorithm follows dark and light edges backward to compute the $h^E$ values shown alongside each node.

$G$ is on the E-graph, though this need not be the case in general. $V^E$ is the pentagon on the right, and its $h^E$ values are precomputed before the search. For $S\notin V^E$, $h^E(S)$ is computed only when A* generates $S$.

@@

We now combine the ideas from HSP2 and E-graphs to construct a domain-independent STRIPS planner which learns from experience. In STRIPS, a consistent heuristic $h(S,S')$ is given by the HSP-max estimate $g_S(S')$. We can compute $g_S(S')$ for all $S,S'\in V^E$ in precisely $|V^E|$ runs of Algorithm \ref{alg:ComputeG}: each run provides us with $|V^E|$ of these results, yielding $(|V^E|)^2$ values in total.

As before, Dijkstra's algorithm combines an inflation of these estimates with the E-graph edges to derive $h^E(S)$ for all $S\in V^E$. Note that in STRIPS, $V^E$ need not include every goal state: the minimal goal $G$ suffices as it's always the easiest to reach in the heuristic relaxation.

Upon encountering a new state $S\notin V^E$, we run Algorithm \ref{alg:ComputeG} again to compute $g_S(a)$ for all atoms $a\in A$. Then it's a simple matter to compute the E-graph heuristic by

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP2 would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which takes O($|V^E||A|$) time. A pseudocode implementation is listed in Algorithm \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while the weight $\epsilon$ provides a uniform goal-directed bias.

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\FORALL{$S \in V^E$}
\STATE ComputeG($S$)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c^E(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated by A*}
\STATE ComputeG($S$)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

For comparison, standard HSP2 without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate each state, and does not incur the cost of generating a state unless the A* frontier encounters it.
When HSP2 is augmented with small E-graphs, the same asymptotic bound continues to hold, provided that the number of E-graph vertices is limited to $O(\mathbf{AO})$. Furthermore, we retain completeness and the ability to configure suboptimality bounds by adjusting the bias parameters:

@@

The previous experimental setup tested our ability to complete partial plans that are fragmented by gaps; however, the experimental searches solved identically the same problems as the control used to generate the E-graph.
Now, we try changing the start and goal states by moving each of them a small distance within the state space. 
In this manner, we test how well E-graphs enable the planner to generalize to nearby but non-identical queries.

>>>>

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
% AAAI format packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
% Additional packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{comment}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
% END Additional packages
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Experience-Biased Symbolic Planning)
/Author (Aram Ebtekar, Mike Phillips, Sven Koenig, Maxim Likhachev)
/Keywords (symbolic planning, E-Graph, experience-biased, weighted A* search, STRIPS, HSP)
}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Experience-Biased Symbolic Planning}
\author{Aram Ebtekar$^\dagger$ \and Mike Phillips$^\dagger$ \and Sven Koenig\thanks{University of Southern California, Los Angeles, CA 90089} \and Maxim Likhachev% <-this % stops a space
\thanks{Carnegie Mellon University, Pittsburgh, PA 15217}% <-this % stops a space
%
}
\author{SoCS 2014 Submission 2}% anonymizer
\maketitle
\begin{abstract}
\begin{quote}
Weighted A* search is the foundation for state-of-the-art domain-independent symbolic planners.
An important open challenge for these planners is to utilize experience with similar planning problems to speed up subsequent planning episodes.
Experience graphs have recently shown promise as a technique for plan reuse in the context of weighted A* searches in robotics, by biasing them towards a subgraph of the search space, typically chosen to consist of the edges used in previous plans.
In this paper, we demonstrate how to augment symbolic planners with
experience graphs, using the HSP2 planner as a representative
example.
The resulting planners are able to trade off seamlessly
between plan quality and runtime and thus have advantages over past
plan-reuse techniques, such as planning by analogy.
We demonstrate experimentally that on many symbolic domains, our approach reduces the number of nodes that the HSP2 planner needs to consider by a factor of 2 or more while achieving about the same plan quality.
\end{quote}
\end{abstract}

@@

In practice, even using state-of-the-art methods to generate heuristics on isolated planning instances, STRIPS problems can be very difficult to solve without expert domain knowledge.
Even humans have difficulty when faced with an unfamiliar kind of puzzle, though we get better with experience.
Thus, one might hope that a planning agent would similarly learn to generalize solutions from past planning experiences to related new instances.
\end{comment}

Modern state-of-the-art symbolic planners tend to rely on weighted A* search as their core algorithm.
In forward weighted A* search, a frontier of nodes expands from the start until it hits a goal, at which point a solution is found.
The performance of weighted A* depends heavily on its \textit{heuristics}, cheaply computable approximations of the remaining distance from a given node to the goal. Heuristics focus the efforts of weighted A* toward the goal and can greatly impact the runtime of the search.
Consequently, much research has been devoted to developing new heuristic functions and different techniques for combining multiple heuristic functions together, all with the aim of providing more effective guidance to the weighted A* search.

However, even with good heuristic functions, many symbolic planning problems can be very difficult to solve.
A state-of-the-art domain-independent planner, or even a human faced with an unfamiliar kind of problem, might have great difficulty in solving it.
After having solved several instances of a problem, humans tend to utilize their experience and learn how to solve similar problems quickly.
Similarly, it is beneficial to develop planning algorithms that can learn to solve repeated planning problems faster over time.

In robotics, Experience Graphs (E-Graphs) have recently been introduced as a means of accelerating motion planning from experience~\cite{phillips2012graphs}.
Based on previously computed plans (a.k.a. experiences), this methods computes an E-Graph that models the high-level connectivity of the free space in motion planning tasks.
The intuition is to remember previously generated paths so that, when a new start and goals are queried, a new path can be quickly generated by reusing subpaths from the E-Graph.
A variant of weighted A* is employed which biases its focus toward the E-Graph.
While this work demonstrates promising results in robot motion planning, E-Graphs have never before been applied to symbolic domain representations such as the STRIPS language.

Our present contribution is to investigate the use of E-Graphs in STRIPS domains. In particular, we take the HSP2 planner~\cite{bonet2001planning} as a representative of the class of modern symbolic planners that are based on weighted A*, and augment its heuristic with a bias toward its E-Graph. After presenting the related works, we begin by describing the STRIPS language and the original HSP heuristic. In the following section, we formally introduce E-Graphs in technical detail. Then we combine the two approaches, resulting in the new Experience-Biased HSP planner. We present the algorithm and discuss its theoretical properties. This is followed by experiments showing promise for the application of E-Graphs to symbolic problem solving. Finally, we conclude with some thoughts on extensions worth investigating.

@@

Classically, case-based reasoning methods 
such as PRIAR~\cite{Kamb:92}, SPA~\cite{Hank:95}, MRL~\cite{Koeh:94}, and NoLimit~\cite{mljournal}
have been used to recall 
previous plans that appear similar to the current problem, and 
then adapt them to meet the new specifications.
Compared to these methods, our work reuses prior experience in
the context of modern weighted A*-based methods for symbolic planning. We also provide
a bound on the suboptimality of the solution.

Recently, there has been work on incorporating prior knowledge 
into modern planners (based on weighted A* and powerful heuristics).
For instance, GPG~\cite{Gere:00} uses planning graphs to adapt a prior path to the 
new scenario by searching locally within windows around plan 
inconsistencies. To guarantee completeness, the windows can grow 
if solutions are not found until they constrain the whole plan.
In this work~\cite{DBLP:conf/iccbr/RosaOB07}, case-based reasoning is used to choose the 
order of state evaluations of EHC (enforced hill-climbing), which 
is a component in modern planners such as FF.

OAKplan~\cite{Serina:2010:KFC:1860143.1860472} is a recent 
case-based reasoning method which converts 
planning instances into a compact graph representation, and 
then uses recent kernel methods for matching labeled graphs to select
a similar case to the problem at hand. The selected case is then 
adapted using LPG-ADAPT~\cite{Fox06planstability}. LPG-ADAPT is a replanning method
that encourages \textit{plan stability}: it tries to make 
sure that the new plan is as similar to the previous plan as possible. 
The method also uses a modern heuristic search approach. While E-Graphs too can encourage stability by reusing old plans, our primary purpose is efficient planning with cost suboptimality bounds, rather than stability.

ERRT-PLAN~\cite{workshop-icaps12-errtplan} presents a very different 
approach to reuse in planning. 
Similar to E-Graphs, ERRT has its roots in motion planning~\cite{Bruce:2002}.
ERRT-PLAN applies this sampling-based motion planner that leverages 
reuse to symbolic planning. The planner randomly chooses to extend 
the search tree toward the goal, toward subgoals from previous 
plans, or to reuse an action from a previous plan. 
The extension operations are based on modern heuristic search
planners.
The stochastic nature of the direction the search progresses toward
allows it to overcome local minima caused by misleading heuristics.

The primary difference between our approach and previous works 
is that we plan with prior experience in a way that comes with provable
bounds on the suboptimality of the solution. The use of E-Graphs also
allows us to formulate this in such a way that the user has control over
this suboptimality bound with an intuitive parameter that trades
reuse (speed) for solution quality.

The HSP2~\cite{bonet2001planning} planner uses the heuristic pioneered by HSP, but replaces greedy hill-climbing with weighted A* search.
We test our approach on HSP2 because the latter is straightforward to describe, and because it is representative of the core techniques that are common to many state-of-the-art domain-independent symbolic planners.
Some newer additions to the field include the causal graph heuristic of the Fast Downward planner~\cite{helmert2006fast} and the landmark heuristic of the LAMA planner~\cite{richter2010lama}. Our approach is orthogonal to these developments and can, in principle, be adapted to work with any search heuristic.

@@

Many of today's state-of-the-art domain-independent STRIPS planners are based on the HSP2 planner~\cite{bonet2001planning}, which we now describe.
It's common to compute heuristics by solving an easier, relaxed version of the original problem. In STRIPS, one might ignore the delete lists of operations.
Since having more atoms makes preconditions and the goal condition more likely to hold, this can only make the problem easier, so the heuristic derived from relaxation never overestimates the true cost.

However, even the relaxed STRIPS planning problem includes set-cover as a special case, making it NP-hard.
Intuitively, we see that the relaxed search space remains exponential in size.
To reduce it, HSP2 decouples the atoms, instead estimating the cost of achieving each individual atom.

Let's say we want a heuristic estimate of the distance from state $S$ to a goal that contains $G$. HSP2 estimates the cost to achieve an atom $a\in A$ from $S$ by $g_S(a) = $
\[\begin{cases} 0  &\mbox{if } a \in S
\\ \min_{op\mid a\in Add(op)} \left(g_S(Prec(op)) + Cost(op)\right)  &\mbox{if } a \notin S \end{cases}\]

The HSP heuristic (also used in HSP2) is then defined by
\[h^{HSP}(S) = g_S(G),\]
the estimated cost of achieving all atoms in $G$. Note that $g_S(\cdot)$ is evaluated on certain atom sets, namely $Prec(op)$ and $G$.
To keep the computations feasible, define $g_S(P) = \max_{p\in P} g_S(p)$ for atom sets $P$.
The resulting HSP-max heuristic is an underestimate; indeed, it satisfies a stronger property called consistency.

\begin{defn} For $\epsilon\ge 1$, a heuristic $h$ is called \textbf{$\epsilon$-consistent} iff $h(S) = 0$ when $S$ is a goal and $h(S) \le \epsilon c(S,S') + h(S')$ for all $S,S'\in V$. $h$ is \textbf{consistent} iff it is $1$-consistent. \end{defn}
\begin{thm} Given an $\epsilon$-consistent heuristic, A* without reexpansions is guaranteed to find a path costing no more than $\epsilon$ times the optimal path cost~\cite{LikGorThr-ara}. \end{thm}

If the atoms were completely independent, a more accurate estimate would be $g_S(P) = \sum_{p\in P} g_S(p)$.
Despite its inconsistency in general, the latter HSP-plus heuristic is often useful in practice, as it uses more state information: it decreases if progress is made towards \textit{any} goal atom, instead of just the most distant goal.

@@

@@

E-Graphs can improve total search time over a series of related queries on the same graph.
Formally, an E-Graph is a subgraph of the state space.
It is updated between queries, typically by adding the solution path from the previous search query.

Suppose we have a consistent heuristic $h(S,S')$ for the distance between arbitrary pairs of states.
That is, $h$ obeys the triangle inequality, is never dominated by an edge cost, and $h(S,S) = 0$ for all $S\in V$.

Planning with E-Graphs is simply a matter of running weighted A* search with the \textbf{E-graph heuristic} $h^E$, derived from $h$.
$h^E$ biases the search to follow E-Graph edges instead of $h$ by penalizing the latter by an inflation factor $\epsilon^E > 1$.
To be precise, define

\[h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N \min \left(\epsilon^E h(S_{i-1},S_i),c^E(S_{i-1},S_i)\right)\]
with the minimization taking place over all goal states $S_N$ and all paths $\pi = \langle S_0,S_1,...,S_N \rangle$ of arbitrary length.

$c^E$ encodes E-Graph edge costs, or $\infty$ if the corresponding edge is not a part of the E-Graph.

In the limit as $\epsilon^E \rightarrow 1$, E-Graph edges offer no advantage because $h(S_{i-1},S_i) \le c^E(S_{i-1},S_i)$. Hence, using the triangle inequality to coalesce the sum,
\[h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N h(S_{i-1},S_i) = \min_{S_N} h(S_0,S_N)\]

Conversely, as $\epsilon^E \rightarrow\infty$, the inflated terms dominate. Hence, the minimizing path becomes one which incurs the least estimated cost outside the E-Graph. In particular, $h^E$ has the following property:

\textit{For sufficiently large $\epsilon^E$, once the A* frontier touches an E-Graph component which includes a goal state, no node outside the E-Graph will ever again need to be expanded.}

This means, for instance, that if the E-Graph already contains a path from the start to a goal, then A*, guided by $h^E$ with large bias parameter $\epsilon^E$, will never leave the E-Graph.
If the E-Graph is small, a solution would promptly be discovered.
In this manner, little work is needed to solve previously encountered subproblems.

We cannot afford to compute $h^E$ according to its literal definition, as there are far too many paths to consider.
Fortunately, there exists a practical means of computing it. By the triangle inequality, any consecutive pair of $h(S_{i-1},S_i)$ terms can be merged into one.
Thus, we lose no generality in restricting $S_i$ to lie on the E-Graph for $0 < i < N$.
The heuristic computation can focus on the E-Graph, using $h(S_{i-1},S_i)$ to estimate the cost of leaving the E-Graph at $S_{i-1}$ before reentering it at $S_i$.

Let $V^E$ be the union of the goal states and the E-Graph's vertices.
In addition to following E-Graph edges, we imagine it's possible to ``jump" from any $S\in V^E$ to any $S'\in V^E$ at cost $\epsilon^E h(S,S')$.
That is, let
\[c'(S,S') = \min\left(c^E(S,S'),~\epsilon^E h(S,S')\right)\]

In a preprocessing stage before the main search, we apply Dijkstra's algorithm once in reverse with the costs $c'$ to compute the estimated distance-to-goal $h^E(S)$ from every $S\in V^E$. Ignoring the computation of $h$ (presently considered as a blackbox), this preprocessing takes $O(|V^E|^2)$ time, which is insignificant for small E-Graphs and goal sets.

Later, when A* generates $S \notin V^E$, we see that we have already precomputed costs of paths consisting of all but the first edge of the minimizing path $\pi$ in the definition of $h^E(S)$. Thus, we derive the mathematically equivalent but much cheaper computation
\[h^E(S) = \min_{S'\in V^E} \left(\epsilon^E h(S,S') + h^E(S')\right)\]

\begin{thm} If the base heuristic $h$ is consistent, then $h^E$ is $\epsilon^E$-consistent~\cite{phillips2012graphs}.\end{thm}

\begin{figure}
	\begin{center}
	\includegraphics[scale=0.5]{Pentagon.png}
	\end{center}
	\caption{An example E-Graph with three edges, and corresponding $h^E$-values. $S$ is not on the E-graph.}
	\label{fig:example}
\end{figure}

Figure \ref{fig:example} illustrates the idea. Suppose the dark edges have cost 3 and make up the E-Graph. Off-E-Graph edges are not shown, as they are only referenced indirectly via $h$; instead, light edges represent the jump estimates $h(S_i,S_j) = 2$ (with one exception: $h(S,G) = 5$), which we inflate by a factor $\epsilon^E=2$ to make 4 (or 10, respectively). Dijkstra's algorithm follows dark and light edges backward to compute the $h^E$ values shown alongside each node.

$G$ is on the E-Graph, though this need not be the case in general. $V^E$ is the pentagon on the right, and its $h^E$ values are precomputed before the search. For $S\notin V^E$, $h^E(S)$ is computed only when A* generates $S$.

@@

We now combine the ideas from HSP2 and E-Graphs to construct a domain-independent STRIPS planner which learns from experience. In STRIPS, a consistent heuristic $h(S,S')$ is given by the HSP-max estimate $g_S(S')$. We can compute $g_S(S')$ for all $S,S'\in V^E$ in precisely $|V^E|$ runs of Algorithm \ref{alg:ComputeG}: each run provides us with $|V^E|$ of these values, yielding $|V^E|^2$ values in total.

As before, Dijkstra's algorithm inflates and combines these estimates with the E-Graph edges to derive $h^E(S)$ for all $S\in V^E$. Note that in STRIPS, $V^E$ need not include every goal state: the minimal goal $G$ suffices as it's always the easiest to reach in the heuristic relaxation.

Upon encountering a new state $S\notin V^E$, we run Algorithm \ref{alg:ComputeG} again to compute $g_S(a)$ for all atoms $a\in A$. Then it's a simple matter to compute the E-Graph heuristic by

@@

We remark that, when E-Graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP2 would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-Graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which takes O($|V^E||A|$) time. A pseudocode implementation is listed in Algorithm \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-Graph, while the weight $\epsilon$ provides a uniform goal-directed bias.

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\FORALL{$S \in V^E$}
\STATE ComputeG($S$)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c^E(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated by A*}
\STATE ComputeG($S$)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-Graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

For comparison, standard HSP2 without E-Graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate each state, and does not incur the cost of generating a state unless the A* frontier encounters it.
When HSP2 is augmented with small E-Graphs, the same asymptotic bound continues to hold, provided that the number of E-Graph vertices is limited to $O(\mathbf{AO})$. Furthermore, we retain completeness and the ability to configure suboptimality bounds by adjusting the bias parameters:

@@

The previous experiment tested our ability to complete partial plans that are fragmented by gaps; however, the experimental searches solved identically the same problems as the control used to generate the E-Graph.
Now, we try changing the start and goal states by moving each of them a small distance within the state space. 
In this manner, we test how well E-Graphs enable the planner to generalize to nearby but non-identical queries.
